\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-09-19 notes/2loopNLSM.tex
added comment about TikZ preamble
Minor edits

\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
\usetikzlibrary{external}  
\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
\tikzexternalize[                     %% Activate externalization
 prefix=images/,                    %% Avoid cluttering the directory
 system call={                       %% Use lualatex in system call
    pdflatex \tikzexternalcheckshellescape -halt-on-error -shell-escape -interaction=batchmode -jobname="\image" "\texsource" || rm "\image.pdf"
 },
]

@@

  We make progress towards understanding how color-dual
  representations must be constructed at the multi-loop level. First
  we identify a cubic theory, \emph{semi-abelian} Yang--Mills, that
  unifies many of the color-dual theories constructed in the
  literature, and provides a construction of color-dual numerators at
  one-loop directly from Feynman rules. Joint consideration of these
  theories highlights that the lifting of one-loop color-kinematics
  dual numerators to two loops may be just as challenging as the step
  from tree level to one loop.  As a first step in understanding the
  two-loop challenge, we construct color-dual two-loop four-point
  non-linear sigma model (NLSM) numerators using generalized unitarity
  methods.  The double-copy of the constructed NLSM numerators with
  themselves, $\mathcal{N}=0$ Yang--Mills, and $\mathcal{N}=4$
  Yang--Mills correctly reproduce the known physical unitarity cuts of
  special Gallileons, Born-Infeld theory, and
  Dirac-Born-Infeld-Volkov-Akulov theory, respectively.  These NLSM
  numerators are in explicit tension with generalizations of the
  one-loop NLSM numerators. \ace{I realize that we all know this, but
    does/will a statement to this effect actually make it into the
    draft?}  Additionally, we revisit the two-loop four-point
  Yang--Mills problem in order to identify the exact point of
  inconsistency in a local color-dual construction.

@@

However, despite the tremendous success in leveraging this duality to
compute gauge and gravity observables to high orders in perturbation
theory \cite{FiveLoopN4, GeneralizedDoubleCopyFiveLoops, Bern:2018jmv,
  Carrasco:2021otn}, color-kinematics remains a conjecture at loop
level. Indeed, there are many examples in the literature where
identifying color-dual representations beyond one-loop has posed a
formidable challenge
\cite{Mogull:2015adi,Johansson:2017bfl,KalinN2TwoLoop} \ace{BDN might
  fit better here rather than the 2loop SQCD}. In this work, we use
the nonlinear sigma model (NLSM) and Yang-Mills, two theories proven
to permit color-dual representations at tree-level
\cite{Feng:2010my,Cachazo:2014xea}, as case studies for advancing our
understanding of the kinematic algebra to the multi-loop level.

@@

With this solver, we are able to compute a two-loop integrand for NLSM
that globally manifests the duality off-shell. This represents a new
benchmark for color-dual representations in non-supersymmetric gauge
theories. However in contrast, we find surprisingly that Yang-Mills
does not permit a globally color-dual representation, \ace{Maybe tone
  down surprise and ref BDN} even when considering the most general
polynomial ansatz of Lorentz covariant kinematics. Concretely, we find
a sharp failure point between the ``bowtie'' cut, and the following
jacobi triple:
\begin{equation}
   \NMCutC
  \Rightarrow
  \JacobiOneDoubleBox +  \JacobiOneCrossedBox + \JacobiOnePentaTriangle \neq 0 \,.
\end{equation}
We comment on the implications of this finding in \cref{conclusions},
and discuss alternative approaches where one must relax field
theoretic locality in order to accommodate the construction of
globally color-dual multi-loop integrands.

\section{Background}\label{background}
To set the stage, in this section we provide an overview of
color-kinematics duality -- both the on-shell and off-shell
constructions. We then detail a color-dual boostrap approach for
constructing integrands in theories for which the off-shell kinematic
algebra has not yet been identified, like Yang-Mills and NLSM.
\subsection{Color-dual Amplitudes}\label{onShellCK}

@@

\subsection{Color-dual Lagrangians}\label{offShellCK}
While the BCJ relations are an on-shell statement about the color-dual
nature of scattering amplitudes, one might aspire to hard code
color-kinematics duality by realizing an off-shell Lagrangian
description of the theory. The double-copy construction of
\cref{DCconstruction} would suggest a manifestly cubic Lagrangian of
the form:
\begin{equation}
\mathcal{L} = P_V^{ab}P_W^{ij}\mathcal{O}^a_i \Box \mathcal{O}^b_j +  V^{abc} W^{ijk}\mathcal{O}^a_i \mathcal{O}^b_j \mathcal{O}^c_k
\end{equation}
where $V^{abc}$ and $W^{ijk}$ are cubic Feynman rules mixing field
operators $\mathcal{O}^{a}_i$ indexed by quantum numbers $a$ and
$i$. We have introduced $P_V^{ab}$ and $P_W^{ij}$ projection operators
to encode non-local kinetic structure that could in principle
participate in the construction. Should the off-shell Feynman rules of
this cubic theory be antisymmetric and satisfy Jacobi identities on
all internal edges,
\begin{align}
\textit{anti-symmetry}:& \quad{}_{a_1}\langle V^{a_2} \rangle_{a_3}+\text{cyc}(23)=0\,,
\\
\textit{Jacobi identity}:& \quad {}_{a_1}\langle V^{a_2} V^{a_3}\rangle_{a_4}+\text{cyc}(234)=0\,,
\end{align}
where 
\begin{equation}
{}_{a_1}\langle V^{a_2}V^{a_3}\cdots V^{a_{n-1}} \rangle_{a_n} \equiv V^{a_1a_2b_1}V^{b_1a_3b_2}\cdots V^{b_{n-3} a_{n-2}a_n}\,.
\end{equation}
This is a sufficient condition of the amplitudes to obey BCJ
relations. \ace{Reword previous 2 sentences} Off-shell descriptions of
color-dual theories are rare, and only a few examples are available in
the literature \cite{Monteiro2011pc, Cheung:2016prv, Cheung:2021zvb,
  Cheung:2020djz, Ben-Shahar:2022ixa, Ben-Shahar:2021zww,
  Ben-Shahar:2021doh}\ace{Are these a reasonable
  collection?  Just grabbed the ones below}. Typically when one goes about constructing color-dual
theories directly from a set of cubic interactions, higher point
Jacobi and antisymmetry constraints require introducing additional
operators. See refs \cite{Ben-Shahar:2022ixa} for some recent
examples. Indeed, as we'll show in \cref{sec:cubic}, Yang-Mills
requires introducing a cubic 2-form interaction in order to manifest
the kinematic algebra just at four-point. Absent information about the
kinematic algebra, one can make progress in the construction of
loop-level color-dual amplitudes by employing a color-dual bootstrap.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
\usetikzlibrary{external}  
\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
\tikzexternalize[                     %% Activate externalization
 prefix=images/,                    %% Avoid cluttering the directory
 system call={                       %% Use lualatex in system call
    pdflatex \tikzexternalcheckshellescape -halt-on-error -shell-escape -interaction=batchmode -jobname="\image" "\texsource" || rm "\image.pdf"
 },
]
%Uncomment this  ^^^^^^^^ to compile in new images in real time. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

  We make progress towards understanding how color-dual
  representations must be constructed at the multi-loop level. First
  we identify a cubic theory, \emph{semi-abelian} Yang--Mills, that
  unifies many of the color-dual theories constructed in the
  literature, and provides a construction of color-dual numerators at
  one-loop directly from Feynman rules. Joint consideration of these
  theories highlights that the lifting of one-loop color-kinematics
  dual numerators to two loops may be just as challenging as the step
  from tree level to one loop.  As a first step in understanding the
  two-loop challenge, we construct color-dual two-loop four-point
  non-linear sigma model (NLSM) numerators using generalized unitarity
  methods.  The double-copy of the constructed NLSM numerators with
  themselves, $\mathcal{N}=0$ Yang--Mills, and $\mathcal{N}=4$
  Yang--Mills correctly reproduce the known physical unitarity cuts of
  special Gallileons, Born-Infeld theory, and
  Dirac-Born-Infeld-Volkov-Akulov theory, respectively.  These NLSM
  numerators are in explicit tension with generalizations of the
  one-loop NLSM numerators. \ace{I realize that we all know this, but
    does/will a statement to this effect actually make it into the
    draft?}
We find that the most general \emph{local} numerator for two-loop four-point numerator for $\mathcal{N}=0$ YM fails global color-kinematics duality and we pinpoint the failure to the kissing-triangles cut.

@@

However, despite the tremendous success in leveraging this duality to
compute gauge and gravity observables to high orders in perturbation
theory \cite{FiveLoopN4, GeneralizedDoubleCopyFiveLoops, Bern:2018jmv,
  Carrasco:2021otn}, color-kinematics remains a conjecture at loop
level. Indeed, there are many examples in the literature where
identifying color-dual representations beyond one-loop has posed a
formidable challenge
\cite{Mogull:2015adi, Johansson:2017bfl, KalinN2TwoLoop, Bern:2015ooa}. In this work, we use
the nonlinear sigma model (NLSM) and Yang-Mills, two theories proven
to permit color-dual representations at tree-level
\cite{Feng:2010my,Cachazo:2014xea}, as case studies for advancing our
understanding of the kinematic algebra to the multi-loop level.

@@

With this solver, we are able to compute a two-loop integrand for NLSM
that globally manifests the duality off-shell. This represents a new
benchmark for color-dual representations in non-supersymmetric gauge
theories.
However in sharp contrast, we find that Yang-Mills
does not permit a globally color-dual representation, even when considering the most general
polynomial ansatz of Lorentz covariant kinematics.\footnote{Globally color-dual local numerators for two-loop pure YM have been studied before but only in the context of a limited ansatz subject to certain loop power counting constraints \cite{Bern:2015ooa}.}
Concretely, the failure point can be pinpointed to a conflict between between the ``bowtie'' cut and the following Jacobi triple
\begin{equation}
   \NMCutC
  \Rightarrow
  \JacobiOneDoubleBox +  \JacobiOneCrossedBox + \JacobiOnePentaTriangle \neq 0 \,.
\end{equation}
We comment on the implications of this finding in \cref{conclusions},
and discuss alternative approaches where one must relax field
theoretic locality in order to accommodate the construction of
globally color-dual multi-loop integrands.

\section{Background}\label{background}
To set the stage, in this section we provide an overview of
color-kinematics duality -- both the on-shell and off-shell
constructions. We then detail a color-dual boostrap approach for
constructing integrands in theories for which the fully off-shell kinematic
algebra has not yet been identified, like Yang-Mills and NLSM.
\subsection{Color-dual Amplitudes}\label{onShellCK}

@@

\subsection{Color-dual Lagrangians}\label{offShellCK}
While the BCJ relations are an on-shell statement about the color-dual
nature of scattering amplitudes, one might aspire to hard code
color-kinematics duality by realizing an off-shell Lagrangian
description of the theory. The double-copy construction of
\cref{DCconstruction} would suggest a manifestly cubic Lagrangian of
the form
\begin{equation}
\mathcal{L} = P_V^{ab}P_W^{ij}\mathcal{O}^a_i \Box \mathcal{O}^b_j +  V^{abc} W^{ijk}\mathcal{O}^a_i \mathcal{O}^b_j \mathcal{O}^c_k
\end{equation}
where $V^{abc}$ and $W^{ijk}$ are cubic Feynman rules mixing field
operators $\mathcal{O}^{a}_i$ indexed by quantum numbers $a$ and
$i$. We have introduced $P_V^{ab}$ and $P_W^{ij}$ projection operators
to encode non-local kinetic structure that could in principle
participate in the construction.
The theory will be color dual if the cubic vertex is antisymmetric and satisfies the Jacobi identity, specifically,
\begin{align}
\textit{anti-symmetry}:& \quad{}_{a_1}\langle V^{a_2} \rangle_{a_3}+\text{cyc}(23)=0
\\
\textit{Jacobi identity}:& \quad {}_{a_1}\langle V^{a_2} V^{a_3}\rangle_{a_4}+\text{cyc}(234)=0
\end{align}
where 
\begin{equation}
{}_{a_1}\langle V^{a_2}V^{a_3}\cdots V^{a_{n-1}} \rangle_{a_n} \equiv V^{a_1a_2b_1}V^{b_1a_3b_2}\cdots V^{b_{n-3} a_{n-2}a_n}\,.
\end{equation}
Off-shell descriptions of
color-dual theories are rare, and only a few examples are available in
the literature \cite{Monteiro2011pc, Cheung:2016prv, Cheung:2021zvb,
  Cheung:2020djz, Ben-Shahar:2022ixa, Ben-Shahar:2021zww,
  Ben-Shahar:2021doh}.
\ace{Are these a reasonable collection?  Just grabbed the ones below}
\jm{These references get repeated in a second.  I'll fix this.}
Typically when one goes about constructing color-dual
theories directly from a set of cubic interactions, higher point
Jacobi and antisymmetry constraints require introducing additional
operators. See refs \cite{Ben-Shahar:2022ixa} for some recent
examples. Indeed, as we'll show in \cref{sec:cubic}, Yang-Mills
requires introducing a cubic 2-form interaction in order to manifest
the kinematic algebra just at four-point. Absent information about the
kinematic algebra, one can make progress in the construction of
loop-level color-dual amplitudes by employing a color-dual bootstrap.
